---
title: "Unemployment Data from FRED"
authors:
  - admin
date: 2020-06-07
summary: "Get and plot unemployment rates from FRED"
categories: ["R Notebook"]
tags: ["R","FRED","Economics","Data Visualization"]
---

# Initialize

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(fredr)
```

# Getting the Data

Getting data official economic data from in R is a breeze. [FRED](https://fred.stlouisfed.org) (the St Louis Federal Reserve) has provided a handy API for getting data like unemployment. 

Making this process even easier is the R package [fredr](https://github.com/sboysel/fredr) which provides convenient wrappers for pulling data.

In order to get this data, you have to first get an API key from FRED. I have stored mine in the `.Renviron` file so I don't accidentally share it when I publish my notebooks. Once you have set it up in your environment, you can use `Sys.getenv()` to set it locally and use it to pull data. 

```{r}
fred_key <- Sys.getenv('FRED_API_KEY')

fredr::fredr_set_key(fred_key)
```

Once you have set up your API key, you can use `purrr::map_dfr` to easily pull mutliple different series by their FRED code.

Here we have pulled four series:

* `UNRATE`: overall unemployment rate
* `UNEMPLOY`: overall unemployment level
* `FLUR`: Florida unemployment rate
* `BOST625URN`: Boston metro area unemployment rate

```{r}
unemployment_data <- map_dfr(c("UNRATE","UNEMPLOY","FLUR","BOST625URN"),fredr::fredr)
```

# Visualize

Now that we have the data, we can visualize.

```{r}
unemployment_data %>% 
  
  # filer data 
  filter(series_id %in% c("UNRATE","FLUR","BOST625URN")) %>% 
  filter(date > as.Date("2007-01-01")) %>% 
  
  # recode and arrange the different series variables
  mutate(series_id = recode(series_id, "BOST625URN" = "Boston",
                            "FLUR" = "Florida",
                            "UNRATE" = "USA")) %>% 
  mutate(series_id = factor(series_id, levels = c("Boston","Florida","USA"))) %>%
  
  # plot the data
  ggplot(aes(x = date, y = value, 
             color = series_id)) +
  
  geom_line(alpha = 0.7, size = 1.4) +
  scale_color_manual(values = c("#00C5CD", "#43CD80", "#EE4000")) +
  ggthemes::theme_fivethirtyeight() +
  labs(color = "", title = "Unemployment Rate") +
  
  # add shaded region for great recession
  annotate("rect", 
           xmin = as.Date("2007-12-01"), 
           xmax = as.Date("2009-06-01"), 
           ymin = 0, 
           ymax = 16,
           alpha = .2)

```

